Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-13 papers/thesis-kirstie/thesis.tex
more updates

\noindent IA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

@@

\[\]
\noindent IB. The Theory of Freezing and Maxwell's Double Tangent Construction

@@

\begin{figure}[h!]
    \centering
     %\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=4cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion. The gray region is where the liquid and the solid coexist, and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz free energy would be for both the liquid and solid phases at a particular temperature and density. Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in. 
%\[\]
\[\]
IC. Classical Density Functional Theory (cDFT)

@@

\[{}\]
II. METHODS

@@

The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which become wider in general as the temperature increases. The radial distance to the peak decreases as the temperature increases, that is, the shells shrink, as does the sphere. The functions can be treated as neglible outside of a distance of about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of the weight function.

\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
III. RESULTS
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]

>>>>

\noindent I. Introduction

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls~\cite{lucretius}. This model, called the Hard-Sphere fluid, has served as 
the primary reference system for fluids. This is due to the fact that the repulsion between atoms is 
much stronger than their attraction, and so the repulsive potential, which is descriptive of hard spheres, 
dominates the interactions between atoms.

%Van der Waal took into account both the attraction, and the respulsion 
%of the impenetrable spheres in devopling his equation of state for an ideal gas in 1xxx.

In the late 1950s, Scaled Particle Theory (SPT) was developed which provided an expression for the 
excess Helmholtz free energy of a homogeneous Hard-Sphere fluid. 
%This .... allowed for ... 
In the 1960s Percus-Yevick applied Density Functional Theory (DFT) to find an expression 
for the excess Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. In the late 1960s, and early 1970s, 
Barker and Henderson adapted the DFT method to use with fluids composed of 
spheres with softly fading potential functions rather than the hard edges of the spheres 
in the Hard-Sphere fluid. This was done by making the radius of the soft spheres temperature dependent, 
so that at low temperatures the fluid could be modeled by hard spheres with large radii, and at high 
temperatures the fluid could be modeled by hard spheres with smaller radii. This is because at higher 
temperatures, the soft-spheres would penetrate into each other farther when they collide. 
Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential, 
which became known as the Weeks-Chandler-Anderson model fluid.

In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT) to form an expression for the excess 
Helmholtz free energy of an inhomogeneous hard sphere fluid based on the results from SPT. In 
19xx Schmidt adopted FMT for soft-sphere atoms to form Soft Fundamental Measure Theory (SFMT). In this paper, 
Density Functional Theory will be used to predict the freezing of a Weeks-Chandler-Anderson model fluid 
by applying SFMT to form the expression for the excess Helmholtz free energy, and the results will be 
compared to those generated from Monte-Carlo simulations.
%\[{}\]
\newline
\newline
\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

@@

\[\]
\noindent IIB. The Theory of Freezing and Maxwell's Double Tangent Construction

@@

\begin{figure}[h!]
    \centering
     %\includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=4cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion. The gray region is where the liquid and the solid coexist, and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz free energy would be for both the liquid and solid phases at a particular temperature and density. Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in. 
%\[\]
\[\]
\noindent IIC. Classical Density Functional Theory (cDFT)

@@

\[{}\]
III. METHODS

@@

The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which become wider in general as the temperature increases. The radial distance to the peak decreases as the temperature increases, that is, the shells shrink, as does the sphere. The functions can be treated as neglible outside of a distance of about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of the weight function.
%\[{}\]

Monte-Carlo integration is employed to more efficiently 
compute the weighted densities used to calculate the excess free energy which is then 
added to the ideal free energy (computed analytically) to get the total free energy 
which is minimized.

\newpage

\noindent IV. Results and Discussion

Phase diagrams for the WCA fluid are shown below. In all cases $\sigma=1$ and $\epsilon=1$.

\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\begin{figure}
 \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
\caption{Phase Diagram of the reduced termperature verses the reduced density. Green lines indicate the region of coexistence from Monte-Carlo simulations.}
\label{fig:Phase_Diagram_of_T_vs_n}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The green line marks the boundary from Monte-Carlo simulations.}
\label{fig:Phase_Diagram_P_vs_T}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
 \end{center}
\caption{Phase Diagram of the reduced temperature verses the reduced density. The grey region indicates coexistence.}
\label{fig:T-vs-n_at_fixed_P}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The lines bending downward indicate freezing.}
\label{fig:p-vs-T_at_fixed_density}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced volume. The grey region indicates coexistence.}
\label{fig:p-vs-V_at_fixed_T}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced density. The grey region indicates coexistence.}
\label{fig:p-vs-n_at_fixed_T}
\end{figure}

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\[{}\]
\[{}\]
\[{}\]
\[{}\]

\noindent V. Conclusion

\[{}\]
\[{}\]
\[{}\]
\[{}\]
